#Logistic Regression

##The Data

*The dataset is imported.*
```{r, echo=TRUE} 
alc <- read.table("alc.txt", header = T, sep = ";")

```
There are **35** variables in the dataset with **382** observations in total. Among the variables there is some basic demographic data (e.g. sex, age, family size, etc.), some variables are related to school, some to personal life and relationships. G1, G2 and G3 are grades for the first period, second period and final grade accordingly.  
  
The variables called *alc_use* shows the average weekly alcohol consumption while the variable *high_use* indicated average weekly alcohol consumption higher than 2.


```{r, echo=TRUE} 
colnames(alc)

str(alc)


```

I will analyse the relationships between alcohol consumption and 4 of the following variables:  
1. weekly study time (*studytime*)  
2. romantic relationship status (*romantic*)  
3. quality of family relationships (*famrel*)  
4. age (*age*)  

##Testing hypothesis  

Four hypothesis related to the alcohol consumption and the chosen variables:  
**H1**: Students who study longer during the week, consume less alcohol compared to those who study less.  
**H2**: Those in a romantic relationship consume less alcohol compared to those who are not in a romantic relationship.  
**H3**: Those who have excellent family relationships consume less alcohol than those in poor family relationships.  
**H4**: Younger students consume more alcohol than older students.  


```{r} 
library(ggplot2)
library(dplyr)

```
###First hypothesis  

The relationship between alcohol consumption and weekly study time will be analysed.  
  
In the survey students had to choose how many hours a week they study. The options were:  
1 - less than 2 hours  
2 - from 2 to 5 hours  
3 - from 5 to 10 hours  
4 - more than 10 hours  
  
Here we see that on average students study from 2 to 5 hours, because the mean of the answers is around 2 which is attributed to 2-10 hours as explained above. In addition to that, all the answers were ranging from 1 and 4.


```{r, echo=TRUE} 
summary(alc$studytime)

```

To visualize the data a boxplot will be drawn.  
  
This grapgh shows that 50% of students who do not consume high amounts of alcohol study from 2 to 10 hours a week. 50% of students who consume high amounts of alcohol study less than 2 hours a week.


```{r, echo=TRUE} 
g1 <- ggplot(alc, aes(x = high_use, y= studytime))
g1 + geom_boxplot(aes()) + xlab("High alcohol consuption") + ylab("Weekly study time") + ggtitle("Student weekly study time by alcohol consumption")

```

The table below shows the frequencies. It is clear that most of the students (195) study from 2 to 5 hours a week, while least students study more than 10 hours. Additionally, overall most students do not consume high amounts of alcohol weekly irrespectivelly of how much they study. However, it is obvious that students that study less than 2 hours a week have a higher proportion of students that consume more alcohol compared to all the other students. 

```{r, echo=TRUE}
table_studytime <- xtabs(~high_use+studytime, data=alc)
ftable(table_studytime)
```
Similar results are visible in the grapgh below.

```{r, echo=TRUE}
counts_studytime <- table(alc$high_use, alc$studytime)
barplot(counts_studytime, main="Study time vs high use",
  xlab="Study time", col=c("darkblue","red"),
 	legend = rownames(counts_studytime), beside=TRUE)
```

**H1**: Students who study longer during the week, consume less alcohol compared to those who study less. 

**H1 is confimed.**  
  
###Second Hypothesis  
  
The relationship between alcohol consumption and relationship status will be analysed. 
  
In the survey students had to choose wheather they are in a romantic relationship (YES) or they aren't (NO).  
  
We can see that more than 2 times more students are not in a romantic relationship.

```{r, echo=TRUE} 
summary(alc$romantic)

```





```{r, echo=TRUE}
mytable <- xtabs(~high_use+romantic, data=alc)
ftable(mytable)
```

```{r, echo=TRUE}
counts <- table(alc$high_use, alc$romantic)
barplot(counts, main="Romantic vs high use",
  xlab="Relationship status", col=c("darkblue","red"),
 	legend = rownames(counts))
```





















```{r, echo=TRUE} 
g2 <- ggplot(alc, aes(x = high_use, y= famrel))
g2 + geom_boxplot(aes()) + xlab("High alcohol consuption") + ylab("Quality of family relationship") + ggtitle("Student quality of family relationship by alcohol consumption")

```

```{r, echo=TRUE} 
g3 <- ggplot(alc, aes(x = high_use, y= freetime))
g3 + geom_boxplot(aes()) + xlab("High alcohol consuption") + ylab("Free time after school") + ggtitle("Student free time after school by alcohol consumption")

```


```{r, echo=TRUE}
mytable <- xtabs(~high_use+romantic, data=alc)
ftable(mytable)
```

```{r, echo=TRUE}
counts <- table(alc$high_use, alc$romantic)
barplot(counts, main="Romantic vs high use",
  xlab="Relationship status", col=c("darkblue","red"),
 	legend = rownames(counts))
```

```{r, echo=TRUE}
counts2 <- table(alc$high_use, alc$freetime)
barplot(counts2, main="Free time vs high use",
  xlab="Free time", col=c("darkblue","red"),
 	legend = rownames(counts))
```

```{r, echo=TRUE}
counts3 <- table(alc$high_use, alc$freetime)
barplot(counts3, main="Free time vs high use",
  xlab="free time", col=c("darkblue","red"),
 	legend = rownames(counts3), beside=TRUE)
```


```{r, echo=TRUE}



```
