#Logistic Regression

##The Data

*The dataset is imported.*
```{r, echo=TRUE} 
alc <- read.table("alc.txt", header = T, sep = ";")

```
There are **35** variables in the dataset with **382** observations in total. Among the variables there is some basic demographic data (e.g. sex, age, family size, etc.), some variables are related to school, some to personal life and relationships. G1, G2 and G3 are grades for the first period, second period and final grade accordingly.  
  
The variables called *alc_use* shows the average weekly alcohol consumption while the variable *high_use* indicated average weekly alcohol consumption higher than 2.


```{r, echo=TRUE} 
colnames(alc)

str(alc)


```

I will analyse the relationships between alcohol consumption and 4 of the following variables:  
1. weekly study time (*studytime*)  
2. free time after school (*freetime*)  
3. quality of family relationships (*famrel*)  
4. age (*age*)  

##Testing hypothesis  

Four hypothesis related to the alcohol consumption and the chosen variables:  
**H1**: Students who study longer during the week, consume less alcohol compared to those who study less.  
**H2**: Those students who have more free time after school consume higher amounts of alcohol than those who have less free time.  
**H3**: Those who have better family relationships consume less alcohol than those in poor family relationships.  
**H4**: Younger students consume more alcohol than older students.  


```{r} 
library(ggplot2)
library(dplyr)

```
###First hypothesis  

The relationship between alcohol consumption and weekly study time will be analysed.  
  
In the survey students had to choose how many hours a week they study. The options were:  
1 - less than 2 hours  
2 - from 2 to 5 hours  
3 - from 5 to 10 hours  
4 - more than 10 hours  
  
Here we see that on average students study from 2 to 5 hours, because the mean of the answers is around 2 which is attributed to 2-10 hours as explained above. In addition to that, all the answers were ranging from 1 and 4.


```{r, echo=TRUE} 
summary(alc$studytime)

```

To visualize the data a boxplot will be drawn.  
  
This grapgh shows that 50% of students who do not consume high amounts of alcohol study from 2 to 10 hours a week. 50% of students who consume high amounts of alcohol study less than 2 hours a week.


```{r, echo=TRUE} 
g1 <- ggplot(alc, aes(x = high_use, y= studytime))
g1 + geom_boxplot(aes()) + xlab("High alcohol consuption") + ylab("Weekly study time") + ggtitle("Student weekly study time by alcohol consumption")

```

The table below shows the frequencies. It is clear that most of the students (195) study from 2 to 5 hours a week, while least students study more than 10 hours. Additionally, overall most students do not consume high amounts of alcohol weekly irrespectivelly of how much they study. However, it is obvious that students that study less than 2 hours a week have a higher proportion of students that consume more alcohol compared to all the other students. 

```{r, echo=TRUE}
table_studytime <- xtabs(~high_use+studytime, data=alc)
ftable(table_studytime)
```
Similar results are visible in the grapgh below.

```{r, echo=TRUE}
counts_studytime <- table(alc$high_use, alc$studytime)
barplot(counts_studytime, main="Study time vs high use",
  xlab="Study time", col=c("darkblue","red"),
 	legend = rownames(counts_studytime), beside=TRUE)
```

**H1**: Students who study longer during the week, consume less alcohol compared to those who study less. 

**H1 is confimed.**  
  
###Second Hypothesis  
  
The relationship between alcohol consumption and amount of free time after school will be analysed. 
  
In the survey students had to choose how much free time after school they have. The possible answers ranged from 1 - very low to 5 - very high.  
  
We can see from the summary below that on average students chose answer 3, which is average amount of free time after school. However, all answers in the range were selected. 

```{r, echo=TRUE} 
summary(alc$freetime)

```

The box plot indicates that 50% of students who consume high amounts of alcohol and 50% of students who don't had the same amount of free time (from average to high).  

```{r, echo=TRUE} 
g4 <- ggplot(alc, aes(x = high_use, y= freetime))
g4 + geom_boxplot(aes()) + xlab("High alcohol consuption") + ylab("Free time after school") + ggtitle("Students' free time after school by alcohol consumption")
```

**H2**: Those students who have more free time after school consume higher amounts of alcohol than those who have less free time.  

**H2 is declined**  


###Third hypothesis  
  
The relationship between alcohol consumption and quality of their family relationship will be analysed.  

In the survey students had to choose wheather what is the quality of their family relationship. The possible answer options ranged from 1 - very bad to 5 - excellent.  
  
On average students indicated that their relationship with the family are good (3.937), however, the whole range of answers was used in answering this question.

```{r, echo=TRUE} 
summary(alc$famrel)

```

Frequencies table also shows that most of the students are in a good relationship with their family. Only 8 students indicated that the relationship was very poor.   

```{r, echo=TRUE}
table_famrel <- xtabs(~high_use+famrel, data=alc)
ftable(table_famrel)
```

The box plot shows that 50% of students who do not consume high amounts of alcohol have good or excellent relationships with their family, whereas, 50% of students who consume high amounts of alcohol have average or good relationships. Another interesting thing is that among the students who do not consume high amounts of alcohol the answers ranged from average to excellent, which outliers for 2 other options. In contast, the answers among the students who consume high amounts of alcohol, the answers ranged from poor to excellent, with an outlier in very poor. Thus, it can be stated that overall students who do not consume high amounts of alcohol are more in better relationships with their families.


```{r, echo=TRUE} 
g2 <- ggplot(alc, aes(x = high_use, y= famrel))
g2 + geom_boxplot(aes()) + xlab("High alcohol consuption") + ylab("Quality of family relationship") + ggtitle("Student quality of family relationship by alcohol consumption")

```

**H3**: Those who have better family relationships consume less alcohol than those in poor family relationships.

**H3 is confirmed.**  

###Fourth hypothesis  
  
The relationship between students' alcohol consumption and their age will be analysed.  
  
In the survey students had to indicate their actual age.  
  
The students' age ranged from 15 to 22 with average age of around 17.

```{r, echo=TRUE} 
summary(alc$age)

```

The boxplot shows that 50% of students who do not consume high amounts of alcohol are between the age of 16 and 17, while 50% of students who consume higher amounts of alcohol are between the ages of 16 and 18 and their overall answer range is wider.  

```{r, echo=TRUE}
g3 <- ggplot(alc, aes(x = high_use, y= age))
g3 + geom_boxplot(aes()) + xlab("High alcohol consuption") + ylab("Age") + ggtitle("Students' age by alcohol consumption")

```


**H4**: Younger students consume more alcohol than older students.

**H4 is declined.**  

##Logistic regression model  

A logistic regression will be used to statistically explore the four variables (*studytime, freetime, famrel, age*) and their relationship with the high/low alcohol consumption variable as the target variable.  
  
**The summary of the logistic regression model**  
1. The standard error shows how much, on average, the estimates would bounce around if the study were re-run indentically, but with new data, over and over.  
2. When the estimate is devided by the standard error, the z value is calculated, which is assumed to be normally distributed.  
3. Pr(>|z|) is the two-tailed p-values that correspond to those z-values in a standard normal distribution.  
4. The significance stars indicate wheather the variables are significant or not. In this case all the variables are statistically significant.  
5. Information about the model and its gooness of fit is in the bottom. The Residual deviance is a measure of the lack of fit of the model taken as a whole, whereas the  Null deviance is such a measure for a reduced model that only includes the intercept.  
6. The null deviance is a model which only contains the intercept. Residual deviance includes the covariants that were chosen and it tests how plausible it is that we would see the data if the model had 377 degrees of freedom. It is comparing the null model (with just the intercept) and checking wheather it's adequate to it. For the the critical value needs to be calculated, which is the chi sq that equals to 423.2743. Since 423 is more than 430, we do reject the null hypothesis and that means that the model is inadequate relative to the null model. In other words it's not better than the null model.  



```{r, echo=TRUE}
m <- glm(high_use ~ studytime + freetime + famrel + age, data = alc, family = "binomial")

summary(m)

#chi sq
qchisq(0.95, 377)
```

The coeficients here mean that:  
If the value of studytime increases by one the model predicts that the log of the odds decreases by 0.55.  
If the value of freetime increases by one the model predicts that the log of the odds increases by 0.35.  
If the value of famrel increases by one the model predicts that the log of the odds decreases by 0.33.  
If the value of age increases by one the model predicts that the log of the odds increases by 0.24.  


```{r, echo=TRUE}
coef(m)
```

Odds ratios for freetime and age are higher than one thus it means they are positivelly associated with high use of alcohol.  
Whereas odd rations for studytime and famrel are lower than 1 so they can't be positivelly associated with higher alcohol consumption.


```{r, echo=TRUE}
OR <- coef(m) %>% exp
CI <- confint(m) %>% exp
cbind(OR, CI)
```

##Predictive Power of the Model  
  
The prediction of high_use being false has a probability of 0.7, while for it beinf true is 0.3.

```{r, echo=TRUE}
probabilities <- predict(m, type = "response")
alc <- mutate(alc, probability = probabilities)
alc <- mutate(alc, prediction = probability > 0.5)
table(high_use = alc$high_use, prediction = alc$prediction)
```
```{r, echo=TRUE}
gprob <- ggplot(alc, aes(x = probability, y = high_use, col = prediction))
gprob + geom_point()
table(high_use = alc$high_use, prediction = alc$prediction) %>% prop.table() %>% addmargins()
```

The average number of wrong predictions in the (training) data is 0.31.

```{r, echo=TRUE}
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}
loss_func(class = alc$high_use, prob = alc$probability)
```

The 10-fold cross-validation on the model has a higher prediction error compared to DataCamp exercise.  

```{r, echo=TRUE}
library(boot)
cv <- cv.glm(data = alc, cost = loss_func, glmfit = m, K = 10)
cv$delta[1]
```
