---
output:
  html_document:
    code_folding: hide
---

#Dimensionality reduction techniques

##The Data

The dataset originates from the United Nations Development Programme. The data has been mutated and now it contains 155 observations and 8 variables in the dataset. The variables in the dataset are listed below. 

```{r, message=FALSE, warning=FALSE}
setwd("C:/Users/Severi/Documents/Open Data/IODS-project/data")
library(MASS)
library(dplyr)
library('corrplot')
library(GGally)
library(tidyr)
library(ggplot2)
human_ <- read.table(file = "human2.txt", sep = ";")
```


```{r}
str(human_)
dim(human_)
```

- *edu2FM* - Female/male ratio in secondary education  
- *labF* - Proportion of females in the labour force  
- *lifeexp* - Life expectancy at birth  
- *eduexp* - Expected years of schooling  
- *gni* -  Gross National Income per capita  
- *matmort* - Maternal mortality ratio  
- *adolbirth* - Adolescent birth rate  
- *repparl* - Percetange of female representatives in parliament  



The graph below shows that variables *edu2FM* (female/male ration in secondary education), *labF* (proportion of females in the labour force), *eduexp* (expected years of schooling) and *repparl* (percentage of female representatives in the parliament) are normally distributed. Correlation seems to be strong between *lifeexp* (life expectancy at birth) and *matmort* (maternal mortality ration); however, to visualize the correlations better, let's look at the next graph.


```{r}
ggpairs(human_)
```

The correlation plot shows strong negative correlation between the previously mentioned variables: *lifeexp* and *matmort*. Also strong negative correlation is between *lifeexp*-*adolbirth*, *eduexp*-*marmort*, and *eduexp*-*adolbirth*. Strong positive correlation is noticable between *lifeexp*-*eduexp* and *marmort*-*adolbirth*.

```{r}
cor_matrix<-cor(human_) %>% round(digits = 2)
corrplot.mixed(cor_matrix)
```

The summaries of the variables are below. The female/male ration in secondary education (*edu2FM*) on average is about 0.85. Proportion of females in the labour fource (*labF*) ranges from 13.5 to 88.1 with average of 52.5. Gross National Income per capita (*gni*) range is really big (from 581 to 123124). Interestingly, our data shows that some Parliaments have no women representatives.

```{r}
summary(human_)
```

##Principal Component Analysis (PCA)

PCA will now be performed on the non-standartized data.The summary below shows that using unstandartized data (with high differences in variable scales) provides us with components that explain very low variance. Thus, the data needs to be standartized.


```{r}
pca_human <- prcomp(human_)
summary(pca_human)
```
The same goes for the biplot below which clearly indicates that *gni* has much bigger scale than all the other variables.

```{r, warning=FALSE}
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```

The data will, therefore, be standartized.  

The summary of variables now shows that the scale is now more similar between different variables, thus the PCA can now be performed again.

```{r}
human_std <- scale(human_)
summary(human_std)
```

The summary shows that now 2 first components explain about 70% of variance, thus the results of the PCA are much better after the data has been standartized.

```{r}
pca_human2 <- prcomp(human_std)
summary(pca_human2)
```

So the biplot now shows high correlation between *matmort* and *adolbirth*. In addition to that the variables *gni*, *eduexp*, *edu2FM* and *lifeexp* seems to be correlated with each other as well. *repparl* is mostly correlated with *labF*; other variables have low/no correlation with *repparl*.

It looks like *matmort*, *adolbirth*, *gni*, *eduexp*, *edu2FM* and *lifeexp* contribute to the dimension of the PC1, while *repparl* and *labF* best define PC2.

The variables that define PC1 are clearly devided into two groups.One group of variables consists of *eduexp* (expected years of schooling), *gni* (Gross National Income per capita), *edu2FM* (female/male ratio in secondary education), and *lifeexp* (life expectancy at birth). The other group defining PC1 contains *matmort* (maternal mortality ratio) and *adolbirth* (adolescent birth rate). These groups together explain education, health and general wealth phenomenons. 

The second principal component that is defined by *repparl* (percentage of female representatives in the parliament) and *labF* (proportion of females in the labour force) explains the gender diversity phenomenon.

The plot indicates that such countries as, e.g. Iceland, Norway, Sweden, Denmank have not only high gross national income per capita but also are high in gender diversity and life expentancy. Whereas, Saudi Arabia, Qatar, Bahrain are high in GNI, but don't have a good situation in gender diversity issues. Countries in Africa seem to have high mortality, poor education and low income.

```{r}
biplot(pca_human2, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```


##Tea dataset

New dataset is now loaded. It is related to tea consumption. The data has 300 observations and 36 variables.

```{r}
library(FactoMineR)
data("tea")
dim(tea)
```
It seems that the variables are related to the time people drink their tea, the places where they drink, how and why they drink it and some other information.

```{r}
str(tea)
```
Only the variables realted to type of tea, way of drinking, frequency of drinking tea and wheather or not tea drinking is feminine or spiritual will be analysed. Their summaries are below.


```{r}
keep_columns <- c("Tea", "How", "how", "feminine", "frequency", "spirituality")
tea_time <- dplyr::select(tea, one_of(keep_columns))
summary(tea_time)
```

The data shows that more respondents think that drinking tea is not a feminine and not a spiritual activity. Most people drink tea more than twice a day. Also more than half of the respondens drink tea using a tea bag. Plain tea without any milk or lemon seems to be the preffered choice, as well as Earl Grey compared to black or green.

```{r, warning=FALSE}
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

There are 12 dimensions based on the summary below.

```{r}
mca <- MCA(tea_time, graph = FALSE)
summary(mca)
```

The MCA factor map shows that people who preffer drinking tea alone, like it to be Earl Grey or green. It seems that those who view tea drinking as non spiritual also think it's not a feminine activity, while those who associate it with femininity also think it's spiritual.

```{r}
plot(mca, invisible = c("ind"), habillage = "quali")
```





