# Regression and model validation

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  



```{r, echo=FALSE} 
students2014 <- read.table("learning2014.txt")

```
The dataset that will be analysed shows that there are 166 observations and 7 variables. Gender is recorded by assigning 1 and 2 for "female" and "male" respectively. Age is recorded as the actual age of the respondents. Questions related to attitude, deep learning, strategic learning and surface learning are evaluated in Likert scale (from 1 to 5). Points are the sum of points that students received from every exam question.

```{r, echo=FALSE}
str(students2014)
```

The dimensions fuction also shows that there are 166 rows (observations) and 7 columns (variables).

```{r, echo=FALSE}
dim(students2014)
```

```{r, echo=FALSE}
library(ggplot2)
p1 <- ggplot(students2014, aes(x = attitude, y = points, col = gender))
p2 <- p1 + geom_point()
p3 <- p2 + geom_smooth(method = "lm")
p4 <- p3 + ggtitle("Student's attitude versus exam points")
p4


```

The gender variable shows that 110 students are female and 56 are male.

```{r, echo=FALSE}
summary(students2014$gender)

```

The average age of the students is 25.51 years old. The youngest respondent is 17 and the oldest 55 y.o. 

```{r, echo=FALSE}
summary(students2014$age)

```

The avarage value for attitude (which was measured in Likert scale (1-5)) is 3.143. Lowest value is 1.4 and the highest value is 5, which was also the highest possible value to choose from.

```{r, echo=FALSE}
summary(students2014$attitude)

```

The avarage value assigned to questions related to deep learning was 3.68. The smallest value was 1.583 and the maximum 4.917.

```{r, echo=FALSE}
summary(students2014$deep)

```

The avarage value of answers to questions related to strategic learning is 3.121. Minimum value is 1.25 and the biggest one 5.

```{r, echo=FALSE}
summary(students2014$stra)

```

The avarage value of answers to questions related to surface learning is 2.787. Minimum value is 1.583 and the biggest one 4.333.

```{r, echo=FALSE}
summary(students2014$surf)

```

The avarage total points that students scored in the exam were 22.72. The lowest score was 7 and the highest 33. The middle value (or the median) was 23, which is almost the same as the average.

```{r, echo=FALSE}
summary(students2014$points)

```





